### 2018-05-22 doc/ms.tex
remove textbf from referee process

%%
%% Beginning of file 'sample61.tex'
%%
%% Modified 2016 September
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.1 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX 
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.1
\documentclass[twocolumn]{aastex61}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex61}
%% 
%% where the arguement options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%  tighten      : Makes baselineskip slightly smaller, only works with 
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for 
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex61}

%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\newcommand{\sm}{M_\odot}
\newcommand{\sr}{R_\odot}
\newcommand{\abc}{iPTF\,16abc}
\newcommand{\sneia}{SNe Ia}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{\today}
\revised{}
\accepted{}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJ}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.1 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{\abc}
\shortauthors{Miller et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
\watermark{DRAFT}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section defines new commands for comments from co-authors
%%
\newcommand{\ycao}[1]{{\color{red} ycao: {#1}}}
\newcommand{\amiller}[1]{{\color{blue} amiller: {#1}}}
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

\title{Early Observations of the Type Ia Supernova \abc:\\
a Case of Interaction with Nearby, Unbound Material and/or Strong Ejecta Mixing}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.1 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affilations. The change was necessary to implement 
%% autoindexing of affilations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.1 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.1 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{A.~A.~Miller}
\email{amiller@northestern.edu}

\author[0000-0001-9515-478X]{A.~A.~Miller}
\affil{Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA) and Department of Physics and Astronomy, Northwestern University, 2145 Sheridan Road, Evanston, IL 60208, USA}
\affil{The Adler Planetarium, Chicago, IL 60605, USA}

\author[0000-0002-8036-8491]{Y.~Cao}
\affil{eScience Institute and Astronomy Department, University of Washington,
  Seattle, WA 98195}

\author[0000-0001-6806-0673]{A.~L.~Piro}
\affil{The Observatories of the Carnegie Institution for Science, 813 Santa Barbara Street, Pasadena, CA 91101, USA}

\author{N.~Blagorodnova}
\affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}

\author{B.~D.~Bue}
\affil{Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA 91109, USA}

\author[0000-0003-1673-970X]{S.~B.~Cenko}
\affil{NASA Goddard Space Flight Center, Mail Code 661, Greenbelt, MD 20771, USA}
\affil{Joint Space-Science Institute, University of Maryland, College Park, MD 20742, USA}

\author{S.~Dhawan}
\affil{The Oskar Klein Centre, Department of Physics, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author{R.~Ferretti}
\affil{The Oskar Klein Centre, Department of Physics, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author[0000-0003-2238-1572]{O.~D.~Fox}
\affil{Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA}

\author{C.~Fremling}
\affil{The Oskar Klein Centre, Department of Astronomy, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author{A.~Goobar}
\affil{The Oskar Klein Centre, Department of Physics, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author[0000-0003-4253-656X]{D.~A.~Howell}
\affil{Las Cumbres Observatory, Goleta, CA 93117, USA}
\affil{Physics Department, University of California, Santa Barbara, CA 93106, USA}

\author[0000-0002-0832-2974]{G.~Hosseinzadeh}
\affil{Las Cumbres Observatory, Goleta, CA 93117, USA}
\affil{Physics Department, University of California, Santa Barbara, CA 93106, USA}

\author[0000-0002-5619-4938]{M.~M.~Kasliwal}
\affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}

\author{R.~R.~Laher}
\affil{Infrared Processing and Analysis Center, California Institute of Technology, Pasadena, CA 91125, USA}

\author{R.~Lunnan}
\affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}

\author{F.~J.~Masci}
\affil{Infrared Processing and Analysis Center, California Institute of Technology, Pasadena, CA 91125, USA}

\author[0000-0001-5807-7893]{C.~McCully}
\affil{Las Cumbres Observatory, Goleta, CA 93117, USA}
\affil{Physics Department, University of California, Santa Barbara, CA 93106, USA}

\author[0000-0002-3389-0586]{P.~E.~Nugent}
\affil{Lawrence Berkeley National Laboratory, Berkeley, California 94720, USA}
\affil{University of California -- Berkeley, Berkeley, CA 94720, USA}

\author{J.~Sollerman}
\affil{The Oskar Klein Centre, Department of Astronomy, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author{F.~Taddia}
\affil{The Oskar Klein Centre, Department of Astronomy, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author[0000-0001-5390-8563]{S. R. Kulkarni}
\affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}

% \author{The Alphabet Friends}
% \affil{the intermediate Palomar Transient Factory}

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.1 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}

Early observations of Type Ia supernovae (SNe) provide a unique probe of their
progenitor systems and explosion physics. Here, we report the intermediate
Palomar Transient Factory (iPTF) discovery of an extraordinarily young SN Ia,
\abc. By fitting a power law to our early light curve, we infer that first
light for the SN, that is when the SN could have first been detected
by our survey, occurred only $0.15\pm_{0.07}^{0.15}$\,d before our
first detection. In the $\sim$24\,hr after discovery, \abc\ rose by
$\sim$2\,mag, featuring a near-linear rise in flux for $\gtrsim$3 d.
Early spectra show strong \ion{C{2} absorption, which disappears
after $\sim$7\,d.} Unlike the extensively-observed Type Ia SN\,2011fe, the
$(B-V)_0$ colors of \abc\ are blue and nearly constant in the days after
explosion. We show that our early observations of \abc\ cannot be explained by
either SN shock breakout and the associated, subsequent cooling, or the SN
ejecta colliding with a stellar companion. Instead, we argue that the early
characteristics of \abc, including: (i) the rapid, near-linear rise, (ii) the
non-evolving blue colors, and (iii) the strong \ion{C{2}} absorption,
are the result of either ejecta interaction with nearby, unbound
material or vigorous mixing of radioactive $^{56}$Ni in the SN ejecta, or a
combination of the two. In the next few years, dozens of very young
\textit{normal} SNe Ia will be discovered, and observations similar to those
presented here will constrain the white dwarf explosion mechanism.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{methods: observational --- surveys --- supernovae: general --- supernovae: individual (\abc, SN\,2011fe)}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.

%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction}
\label{sec:intro}

Although Type Ia supernovae (SNe Ia) have been extensively used as
standardizable candles, their progenitor systems and explosion physics are
still debated (see a recent review by \citealt{2014ARA&A..52..107M}).
Extremely detailed observations in the hours to days after explosion provide a
promising avenue to further constrain this problem.

While the shock breakout of a SN Ia occurs on a sub-second timescale, the
subsequent quasi-adiabatic expansion and cooling of the unbound ejecta
produces thermal emission that can be used to infer the radius of the
exploding star \citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing
models of this cooling emission to the earliest-phase data of SN~2011fe,
\citet{2012ApJ...744L..17B} concluded that the explosion came from a star with
$R_\ast \lesssim 0.02\;\sr$, where $\sr$ is the solar radius. Combining the
radius constraint with the measured ejecta mass,
\citeauthor{2012ApJ...744L..17B} derive the mean density of the progenitor
star, confirming that at least some Type Ia SNe come from compact and
degenerate stars.

Early phase observations of SNe Ia from a white dwarf (WD)$+$non-degenerate
binary may detect excess emission, relative to most Type Ia SNe, due to the
collision of the SN ejecta with the non-degenerate companion
\citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess emission was
first detected in iPTF\,14atg \citep{2015Natur.521..328C}, a low-velocity SN
Ia with a significant and declining ultraviolet (UV) pulse detected within a
few days of the SN explosion. This UV pulse is best interpreted as a SN
ejecta-companion collision (but see also
\citealt{2016MNRAS.459.4428K,2017MNRAS.472.2787N}). While such emission
requires a favorable geometric alignment and is only expected in
$\lesssim$10\% of SNe Ia \citep{2010ApJ...708.1025K}, many studies have
searched for signatures of an ejecta-companion interaction, typically
resulting in non-detections (e.g.,
\citealt{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
2012ApJ...744L..17B,2015Natur.521..332O,
2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
2015ApJS..221...22I}). Possible exceptions include SN\,2012cg, which exhibited
excess blue emission in its early-phase light curve
(\citealt{2016ApJ...820...92M}; though see \citealt{2016arXiv161007601S} for
an interpretation that does not invoke ejecta-companion interaction), and
SN\,2017cbv, which shows a clearly resolved ``bump'' in the early $UBg$ light
curves \citep{2017ApJ...845L..11H}.
% With only $\sim 10\%$ of single-degenerate progenitors occupying the preferred binary geometry for us to see the collision
% signatures, it is not surprising that there are few confirmed detections in the literature \ycao{ref?}.

Interaction is not limited to systems with a non-degenerate companion,
however, as WDs enshrouded in diffuse material following a binary merger
(e.g., \citealt{2015MNRAS.447.2803L}) or expanded due to a pre-explosion
pulsation can give rise to ejecta-interaction signatures (e.g.,
\citealt{2014MNRAS.441..532D}). Models of this scenario naturally produce
\ion{C}{2} absorption that is comparable in strength to \ion{Si}{2} in the
days after explosion \citep{2014MNRAS.441..532D}, as was observed in
SN\,2013dy \citep{2013ApJ...778L..15Z} and SN\,2017cbv
\citep{2017ApJ...845L..11H}.

The vast majority of SNe Ia are observed to be powered purely by the
radioactive decay of $^{56}$Ni. While the detection of SN shock cooling or
ejecta interaction is rare, the level of $^{56}$Ni mixing in the SN ejecta can
fundamentally alter the appearance of the SN shortly after explosion
(e.g.,
\citealt{2014MNRAS.441..532D,2016ApJ...826...96P,2017MNRAS.472.2787N)}.

SNe Ia experience a dark phase after the SN shock breakout but before
radioactive energy diffuses into the photosphere \citep{2014ApJ...784...85P}.
The duration of this dark phase is set by how the newly synthesized $^{56}$Ni
is mixed and deposited into different layers of the ejecta. Strong
mixing leads to a short, or non-existent, dark phase, because the radioactive
$\gamma$-rays rapidly diffuse to the photosphere. This also leads to
larger luminosities and bluer optical colors at early times.
Even with vigorous mixing it is difficult at very early times, $\ll
1$\,d after explosion, to explain very large luminosities or blue colors,
because the $^{56}$Ni has not had sufficient time to radioactively decay to
$^{56}$Co.} If the mixing is weak and the $^{56}$Ni is confined to the
innermost layers of the ejecta, the dark phase can last several days.
Weak mixing results in redder colors and a more moderate rise in
luminosity \citep{2014MNRAS.441..532D,2016ApJ...826...96P}. Thus, the early
light curves of even non-exotic Type Ia SNe convey information about their
progenitor systems and explosion mechanisms by constraining the distribution
of $^{56}$Ni.
% \footnote{\citet{2015A&A...574A..60T} found that a significant dark
% phase is not prevalent for a sample of Type Ib/c SNe.}

\citet{2017MNRAS.472.2787N} demonstrate that disambiguating between these
different scenarios via optical photometry alone is challenging.
\citeauthor{2017MNRAS.472.2787N} further show that estimates of the time of
explosion, which are critical for comparing models with observations, are
often incorrect by as much as $\sim$2\,d using common methods in the
literature. While analytical models suggest that early spectra can be used to
infer the time of explosion (e.g., \citealt{2014ApJ...784...85P}), more
detailed simulations show that the photospheric evolution is not so simple
\citep{2016ApJ...826...96P}. Reconciling these issues requires both a larger
sample of early SNe Ia observations and more detailed models that produce
synthetic light curves and spectra.

In this paper, we report observations of an extraordinarily young SN Ia, \abc,
which was discovered by the intermediate Palomar Transient Factory (iPTF) on
2016 April $3.36$ UTC at $\textrm{R.A.}=13^h34^m45.49^s$,
$\textrm{Dec.}=+13\degr51\arcmin14\farcs3$ (J2000) with a
$g_\mathrm{PTF}$-band magnitude of $21.44\pm0.25$ \citep{2016ATel.8907....1M}.
The transient is spatially coincident with a tidal tail of the galaxy
NGC\,5221, which lies at a distance of $\sim$100\,Mpc. \abc\ is not detected
to a 5-$\sigma$ limit of $g_\mathrm{PTF}=21.9$\,mag on April $2.42$,
less than 1\,d prior to discovery, and rose by $\sim$2\,mag in the 24\,hr
following its initial detection. Our spectroscopic follow-up campaign
classified \abc\ as a normal SN Ia \citep{2016ATel.8909....1C}. Our
observations and analysis show that the early evolution of \abc\ exhibited
several distinct properties relative to SN\,2011fe. We interpret those
differences as arising from either strong $^{56}$Ni mixing or ejecta
interaction with nearby, unbound material, or a combination of the
two. Alongside this paper, we have released our open source analysis and all
of the data utilized in this study. These are available online at
\texttt{GitHub} \url{https://github.com/adamamiller/iPTF16abc}.

% This paper is organized as follows: Section \ref{sec:obs} describes
% photometric and spectroscopic observations of \abc. Section
% \ref{sec:usual_staff} establishes that \abc\ is a normal SN Ia
% in NGC\,5221. Section \ref{sec:first_light} analyzes the early
% light curve and spectra.

\section{Observations}
\label{sec:obs}

During the spring of 2016, the iPTF survey observed the field of \abc\ every
night during dark time in either the $g_\mathrm{PTF}$ or $R_\mathrm{PTF}$
band.\footnote{P48 observations of \abc\ are reported in the $g_\mathrm{PTF}$
and $R_\mathrm{PTF}$ filters throughout, which are similar to the SDSS $g'$
and Mould-$R$ filters, respectively (see \citealt{2012PASP..124..854O} for
details on PTF calibration). The correction from the $g_\mathrm{PTF}$ and
$R_\mathrm{PTF}$ filters to SDSS $g'$ and $r'$ requires knowledge of the
intrinsic source color (see Eqns.~1 and 2 in \citealt{2012PASP..124..854O}).
The spectral diversity of SNe Ia in the days after explosion is poorly
constrained, and as a result the color terms for \abc\ at these epochs are
unknown. We proceed by assuming the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$
calibration is on the AB system, which strictly speaking is incorrect, but
this does not fundamentally alter any of our conclusions.} Survey observations
were conducted with the CFH12K camera \citep{2008SPIE.7014E..4YR} on the
Palomar Observatory 48-inch telescope (P48; \citealt{2009PASP..121.1395L}).
Images were processed by the IPAC image-subtraction pipeline, which subtracts
background galaxy light using deep pre-SN images, and performs forced
point-spread-function (PSF) photometry at the location of the SN
\citep{2017PASP..129a4002M}. The photometry is then calibrated to the PTF
photometric catalog \citep{2012PASP..124..854O}.

\begin{figure}[htb] 
    \centering
    \includegraphics[width=3.35in]{logLC_with_colors.pdf} 
    \caption{
    %
    The $g$-band light curve of \abc, with 5-$\sigma$ upper limits shown as
    downward arrows. Observations from different telescopes are shown with
    different symbols. The lower axis shows time measured in
    rest-frame days relative to $t_0$ (see \S\ref{sec:lc_fit}), while
    the upper axis shows time relative to $B$-band maximum. Note that the
    horizontal axis is shown with a linear scale for $-2 \; \mathrm{d} \le t -
    t_0 \le 3 \; \mathrm{d}$ and a $\log$ scale for $t - t_0 > 3 \;
    \mathrm{d}$. Vertical black ticks show epochs of spectroscopic
    observations. }
% 
    \label{fig:lightcurve} 
\end{figure}

After discovery, $g'$, $r'$ and $i'$-band photometry was obtained
with the SED Machine (SEDm; \citealt{2017arXiv171002917B}) mounted on the
Palomar Observatory 60-inch telescope (P60). We utilized the Fremling
Automated Pipeline (\texttt{FPipe}; \citealt{2016A&A...593A..68F}) to subtract
galaxy light from the SEDm images using archival Sloan Digital Sky Survey
(SDSS) images as a reference. This pipeline then performed forced-PSF
photometry at the location of \abc, which is calibrated to the SDSS catalog
\citep{2014ApJS..211...17A}.

The Las Cumbres Observatory (LCO) 1-m telescope network obtained
$BVg'r'i'$ photometry. PSF photometry was measured on these images using the
\texttt{lcogtsnpipe} pipeline \citep{2016MNRAS.459.3939V}. The \textit{BV}
magnitudes are calibrated to the Fourth USNO CCD Astrograph Catalog
\citep{2013AJ....145...44Z}, and the \textit{g'r'i'} magnitudes are calibrated
to SDSS Data Release 6 \citep{2008ApJS..175..297A}.

The Reionization and Transients InfraRed (RATIR) camera on the autonomous
1.5\,m Harold L. Johnson Telescope
\citep{2012SPIE.8446E..10B,2012SPIE.8444E..5LW} was used to observe \abc\ in
the $r'i'ZYJH$ filters. By design, RATIR lacks a cold shutter, which means IR
dark frames are not available. Laboratory testing, however, confirms that the
dark current is negligible in both IR detectors \citep{2012SPIE.8453E..1OF}.

The RATIR data were reduced, coadded, and analyzed using standard CCD and IR
processing techniques in \texttt{IDL}, \texttt{Python}, \texttt{SExtractor}
\citep{1996A&AS..117..393B}, and \texttt{SWarp}. Aperture photometry is
obtained following the methods described in \citet{2014AJ....148....2L}. The
$r'i'Z$ filters are calibrated to SDSS \citep{2014ApJS..211...17A}, while the
$JH$ filters are calibrated to the Two Micron All Sky Survey
\citep{2006AJ....131.1163S}. For the $Y$-band calibration, we used an
empirical relation in terms of the $J$ and $H$ magnitudes derived from the
United Kingdom Infrared Telescope (UKIRT; \citealt{2007A&A...467..777C}) Wide
Field Camera observations \citep{2009MNRAS.394..675H}.

The \textit{Swift} satellite observed \abc\ on 14 epochs, beginning
$\sim$15\,d pre-maximum light through $\sim$22\,d post maximum. The SN flux is
measured via aperture photometry on Ultraviolet-Optical Telescope (UVOT)
images via the usual procedures in \texttt{HEASoft}, including corrections for
coincidence loss and aperture loss. The image counts are converted to physical
fluxes using the latest calibration \citep{2011AIPC.1358..373B}. There are no
pre-SN UVOT images at the SN location in the \textit{Swift} archive. Visual
inspection of the UVOT images suggests negligible host-galaxy contamination in
our UVOT flux measurements. No X-ray emission is detected from \abc\ by the
\textit{Swift} X-ray Telescope (XRT).

The $g$-band discovery and follow-up data of \abc\ are illustrated in
Figure~\ref{fig:lightcurve}. The photometry is shown in the AB system. As
previously noted the color terms necessary to convert $g_\mathrm{PTF}$ to the
AB system are unknown and assumed to be zero.

Spectroscopic observations of \abc\ were taken with a variety of telescopes
and instruments over multiple epochs beginning $\sim$2\,d after
discovery and ending $\sim$2 months after $B$-band maximum. An
observing log is listed in Table \ref{tab:spec_obs_log}. The spectra were
reduced using standard procedures in
\texttt{IDL}/\texttt{Python}/\texttt{Matlab}. The optical spectral evolution
of \abc\ is illustrated in Figure \ref{fig:spec_seq}, which excludes
high-resolution Very Large Telescope (VLT) spectra for clarity.

\begin{deluxetable}{cccccc}
  \tablecaption{Spectroscopic observations of \abc\ \label{tab:spec_obs_log}}
  \tabletypesize{\footnotesize}
  \tablehead{
    \colhead{Observation} & \colhead{SN} & \colhead{} &
    \colhead{} & \colhead{Range} \\
    \colhead{MJD} & \colhead{phase} & \colhead{Telescope} &
    \colhead{Instrument} & \colhead{(\AA)}
  }
  \startdata
  $57483.26$ & $-15.9$ & DCT & DeVeny\tablenotemark{1} & $3301$--$7499$ \\
  $57483.88$ & $-15.3$ & Gemini-North & GMOS\tablenotemark{2} & $3800$--$9200$ \\
  $57484.51$ & $-14.7$ & Keck-II & DEIMOS\tablenotemark{3} & $5500$--$8099$ \\
  $57486.51$ & $-12.7$ & Keck-II & DEIMOS\tablenotemark{3} & $5500$--$8099$ \\
  $57488.38$ & $-10.9$ & Keck-I & LRIS\tablenotemark{4} & $3055$--$10411$ \\
  $57489.51$ & $ -9.8$ & LCO-2m & FLOYDS\tablenotemark{5} & $3301$--$8999$ \\
  $57490.40$ & $ -8.9$ & LCO-2m & FLOYDS\tablenotemark{5} & $3301$--$9999$ \\
  $57491.55$ & $ -7.8$ & LCO-2m & FLOYDS\tablenotemark{5} & $3300$--$9998$ \\
  $57492.20$ & $ -7.2$ & VLT & X-shooter\tablenotemark{6} & $3300$--$24550$ \\
  $57494.00$ & $ -5.4$ & VLT & UVES\tablenotemark{7} & \\
  $57503.32$ & $ +3.7$ & LCO-2m & FLOYDS\tablenotemark{5} & $3300$--$9999$ \\
  $57506.00$ & $ +6.3$ & NOT & ALFOSC\tablenotemark{8} & $3602$--$8098$ \\
  $57508.27$ & $ +8.5$ & LCO-2m & FLOYDS\tablenotemark{5} & $3301$--$9999$ \\
  $57518.42$ & $+18.5$ & Keck-I & LRIS\tablenotemark{4} & $3071$--$10208$ \\
  $57520.03$ & $+20.0$ & VLT & X-shooter\tablenotemark{6} & $3300$--$24789$ \\
  $57529.40$ & $+29.2$ & LCO-2m & FLOYDS\tablenotemark{5} & $4000$--$8998$ \\
  $57542.41$ & $+41.9$ & LCO-2m & FLOYDS\tablenotemark{5} & $4000$--$8998$ \\
  $57550.40$ & $+49.7$ & LCO-2m & FLOYDS\tablenotemark{5} & $4001$--$8999$ \\
  $57562.38$ & $+61.4$ & LCO-2m & FLOYDS\tablenotemark{5} & $4800$--$9300$ \\
  \enddata
  \tablenotetext{1}{The Deveny Spectrograph \citep{2014SPIE.9147E..2NB}}
  \tablenotetext{2}{The Gemini Multi-Object Spectrograph \citep{2004PASP..116..425H}}
  \tablenotetext{3}{DEep Imaging Multi-Object Spectrograph \citep{2003SPIE.4841.1657F}}
  \tablenotetext{4}{Low-Resolution Imaging Spectrometer \citep{1995PASP..107..375O}}
  \tablenotetext{5}{FLOYDS \url{https://lco.global/observatory/instruments/floyds}}
  \tablenotetext{6}{X-shooter \citep{XShooter}}
  \tablenotetext{7}{Ultraviolet and Visual Echelle Spectrograph \citep{2000SPIE.4008..534D}}
  \tablenotetext{8}{The Andalucia Faint Object Spectrograph and Camera \url{http://www.not.iac.es/instruments/alfosc}}
\end{deluxetable}

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=5.5in]{spectra.pdf}
  \caption{
  Observed spectral sequence of \abc. The spectra are normalized by their
  median flux between 6000 and 7000$\,\textrm{\AA}$. The phase of each
  spectrum relative to the time of $B$-band maximum is shown.
  Telluric absorption bands are grayed out. Line identifications are provided
  for the spectral features discussed in the text. For clarity,
  high-resolution spectra obtained with the VLT have been omitted (see
  \citealt{2017A&A...606A.111F} for a detailed discussion of these spectra).}
  \label{fig:spec_seq}
\end{figure*}

\section{Host Galaxy, Reddening, and Classification}
\label{sec:usual_staff}

\subsection{Host Galaxy}
\label{sec:host}

\abc\ is spatially coincident with a tidal tail of galaxy NGC\,5221.
\citet{2007A&A...465...71T} derived a distance modulus of
$35.0\pm0.4\,\textrm{mag}$ to NGC\,5221 from the Tully-Fisher relation,
consistent with our derivation from the SN light curve (see
\S\ref{sec:classification}).

Separately, \citet{2015MNRAS.447.1531C} observe the 21-cm line in NGC\,5221
and measure a redshift of $0.0234$, which we adopt for the remaining analysis
in this paper.

\subsection{Reddening}
\label{sec:reddening}

A detailed study of the reddening towards \abc\ is presented in a companion
paper \citep{2017A&A...606A.111F}. Briefly, the foreground Galactic extinction
toward \abc\ is $E(B-V) = 0.0279$\,mag \citep{2011ApJ...737..103S}.
High-resolution spectra of \abc\ show multiple absorption components for both
the \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets. Despite large
equivalent widths (EWs) for these lines, implying significant extinction
(e.g., \citealt{2012MNRAS.426.1465P), \citeauthor{2017A&A...606A.111F}~find
evidence for only a small amount of extintion.} The empirical relation between
the EW of \ion{Na}{1}\,D and extinction has a large scatter, and
\citet{2013ApJ...779...38P} have shown that \ion{Na}{1}\,D absorption is a
poor tracer of reddening in Type Ia SNe. Thus, we adopt $E(B-V) = 0.05 \,
\mathrm{mag}$ as the local extinction for \abc\ \citep{2017A&A...606A.111F}.
For the remainder of our analysis we assume a total, Galactic $+$ host galaxy,
line-of-sight extinction of $E(B-V) = 0.08 \, \mathrm{mag}$.

\subsection{Classification}
\label{sec:classification}

Using the SuperNova IDentification (\texttt{SNID};
\citealt{2007ApJ...666.1024B}) package, we find the low-resolution spectrum of
\abc\ at $+18.8$\,d is best matched by normal SNe Ia. Several characteristic
features of a SN Ia, such as \ion{Si}{2} and \ion{S}{2}, can be easily
identified in \abc\ (Figure \ref{fig:spec_seq}). From the $+3.7 \, \mathrm{d}$
LCO spectrum, we measure the pseudo-equivalent widths (pEWs) of $-12 \pm
2$\,\AA\ and $-55 \pm 5$\,\AA\ for the absorption features near 5750 and 6100
\AA, attributed to \ion{Si}{2}\,$\lambda\lambda$5972, 6355, respectively.
According to the \citet{2006PASP..118..560B} classification scheme, \abc\ is a
shallow silicon SN, similar to 1999aa-like SNe \citep{2009PASP..121..238B}.
In Figure~\ref{fig:branch_vel, the velocity evolution of the \abc\
\ion{Si}{2}$\,\lambda6355$ absorption minimum is compared to the median
evolution of the four spectroscopic subclasses from
\citet{2006PASP..118..560B}. The median evolution is defined using the sample
of SNe Ia in \citet{2012AJ....143..126B}. For each SN, we interpolate the
\ion{Si}{2} velocity, $v_{\mathrm{Si\,II}\,\lambda6355}$, to a fixed grid at
1\,d intervals. The curves are defined by the median
$v_{\mathrm{Si\,II}\,\lambda6355}$ at each point on the grid with at least 3
SNe (this prevents just 1 or 2 SNe from defining the evolution of the entire
subclass). The velocity evolution of \abc\ is most reminicent of the
shallow-silicon subclass.}

\begin{figure}[htb]
  \centering
  \includegraphics[width=3.35in]{Branch_vel_evol.pdf}
  \caption{
  Velocity evolution of \ion{Si{2},
  $v_{\mathrm{Si\,II}\,\lambda6355}$, for \abc\ compared to the median
  evolution of the four spectroscopic subclasses defined in
  \citet{2006PASP..118..560B}, using data from \citet{2012AJ....143..126B}.
  Typical uncertainties for \abc\ are $\sim$1000\,km\,s$^{-1}$ before
  $T_{B,\mathrm{max}}$, when the \ion{Si}{2}$\,\lambda6355$ profile is shallow
  and the minimum of absorption is difficult to determine, and
  $\sim$300\,km\,s$^{-1}$ after $T_{B,\mathrm{max}}$. For the median curves,
  the typical scatter, determined via the interquartile range of the sample,
  is $\sim$700\,km\,s$^{-1}$, around $T_{B,\mathrm{max}}$. At early times,
  core-normal and broad-line SNe have significantly faster \ion{Si}{2} than
  \abc, while the declining trend of cool SNe does not match \abc.} }
  \label{fig:branch_vel}
\end{figure}

To determine the brightness and time of $B$-band maximum for \abc, we fit the
P60 light curves with the \texttt{sncosmo} software
package.\footnote{\texttt{sncosmo} is available at
\url{https://sncosmo.readthedocs.io}.} This fit includes a \texttt{SALT2}
template \citep{2007A&A...466...11G} that has been corrected for extinction
using the \citet{1999PASP..111...63F} reddening law, $R_V=3.1$, and $E(B-V) =
0.08 \, \mathrm{mag}$.

We determine the time of rest-frame \textit{B}-band maximum to be
$\textrm{MJD}_\mathrm{max}=57499.54\pm0.23$, the coefficient of the zeroth
principle component $x_0 = 0.0086 \pm 0.0003$, the coefficient of the first
principle component $x_1 = 0.96 \pm 0.15$, and the color term $c = 0.033 \pm
0.029$. The best-fit model also gives an unreddened apparent peak magnitude of
$m^*_{B}=15.80 \pm 0.04 \,\textrm{mag}$ in the SN rest frame. In the following
sections, we adopt $\textrm{MJD}_\mathrm{max}=57499.54$ as the time of
$B$-band maximum, $T_{B,_\mathrm{max}}$, and phase $t=0$.

We measure the (pseudo)bolometric luminosity, $L_{UVOIR}$, of \abc\ at peak
via trapezoidal integration of the reddening-corrected flux from the UV,
optical, and NIR (UVOIR) filters. The light curves in the individual filters
are interpolated so that $L_{UVOIR}$ is evaluated at common epochs in each
filter. From this integration, we measure a maximum luminosity $L_\mathrm{max}
= 1.2 \pm 0.1 \times 10^{43} \; \mathrm{erg \; s}^{-1}$ for \abc. This value
is consistent with the normal SNe Ia studied in \citet{2016A&A...588A..84D}.
Following Arnett's rule \citep{1982ApJ...253..785A,
1985Natur.314..337A}, the mass of $^{56}$Ni synthesized in the explosion can
be derived from $L_\mathrm{max}$. Assuming a rise time of $19 \pm 3$\,d (see
\citealt{2006A&A...450..241S}) we find $M_\mathrm{Ni} = 0.6 \pm 0.1
M_\odot$.\footnote{A 17.9\,d rise time (\S\ref{sec:lc_fit}),
yields a consistent estimate of $M_\mathrm{Ni}$.}}

After establishing \abc\ as a normal SN Ia, we use the latest calibration
\citep{2014A&A...568A..22B} of the Phillips relation
\citep{1993ApJ...413L.105P} using $m^*_{B}$, $x_1$ and $c$ to derive a
distance modulus $\mu = 34.89 \pm 0.10 \,\textrm{mag}$ to the SN, provided
that the host galaxy of \abc\ has a stellar mass $< 10^{10}\sm$. A more
massive host galaxy would result in a larger inferred distance modulus that is
nevertheless consistent within the uncertainties. For the following analysis
we adopt a distance modulus $\mu = 34.89 \pm 0.10 \,\textrm{mag}$ for
\abc.\footnote{This $\mu$ is consistent with the $z_\mathrm{SN}$, $H_0 = 73 \,
\mathrm{km \, s^{-1} \, Mpc}^{-1}$, and Virgo-infall-corrected distance
\citep{2000ApJ...529..786M}.}

\section{Early Observations}
\label{sec:first_light}

Here we consider our suite of early observations of \abc, and compare our
findings with SN\,2011fe, a well-studied, nearby SN that was discovered
shortly after explosion
\citep{2011Natur.480..344N,2012ApJ...744L..17B,2014ApJ...784...85P}.

\subsection{Time of First Light from the Early Light Curve}
\label{sec:lc_fit}

The time of first light for SNe is usually estimated by extrapolating
early-phase light curves to determine when the SN flux is equal to 0.
Assuming an ideal, expanding fireball with constant temperature,
\citet{1982ApJ...253..785A} derives that $f \propto t^2$, where $f$ is the SN
flux and $t$ is the time since explosion. Despite these simplified assumptions, multiple studies have found that the
early emission from Type Ia SNe can be described as a power law in time, with
power-law index consistent with 2, i.e. $f \propto t^2$ (e.g.,
\citealt{2006AJ....132.1707C, 2010ApJ...712..350H,
2011MNRAS.416.2607G}).\footnote{Many of these studies
sample SNe Ia light curves at phases closer to $T_{B,\mathrm{max}}$
than our initial observations of \abc.}

As our observations include especially early observations of \abc, the SN is
not detected $\sim$1\,d prior to discovery, we model the early flux from \abc\
as a power law:

\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t \le t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.\ ,
\end{equation}
% 
where $t_0$ is the time of first light, $\alpha$ is the power-law index, and
$t$ is measured in the SN rest-frame. We allow $\alpha$ to vary to
find the best match to the data, and later show that $\alpha = 2$ is not
compatible with the observations. To determine $t_0$ and $\alpha$ we fit the
earliest observations of \abc. Due to slight variations in the passbands, the
model is fit only to the relative $g_\mathrm{PTF}$-band flux. $g_\mathrm{PTF}$
is the only filter with observations prior to first light, a necessity for
constraining $t_0$.

To determine the best fit parameters, we search a large grid over $t_0$,
$\alpha$, and the proportionality constant, and minimize $\chi^2$. The
modeling results show that the SN flux rises approximately linearly between
$t=-18\,\mathrm{d}$ and $t=-15\,\mathrm{d}$. Figure \ref{fig:early_lc_fit}
shows the best-fit result and the joint marginal distribution of $t_0$ and
$\alpha$. From the best-fit model we obtain $\alpha=0.98 \pm ^{0.16}_{0.14}$
and $t_0=-17.91 \pm ^{0.07}_{0.15}$ d, where the uncertainties represent the
marginalized 95\% confidence intervals. Our first detection of \abc\ occurred
$\sim$${0.15}\,\textrm{d}$ after $t_0$. In the analysis that
follows, the precise values of the best-fit parameters are not important. The
critical finding here is that $\alpha \approx 1$ and $t_0 \approx -18 \,
\mathrm{d}$.

Figure \ref{fig:early_lc_fit} also shows the best-fit model while fixing
$\alpha = 2$. The $f \propto t^2$ model does not match the observations.
Formally, for the $\alpha = 2$ model $\chi^2 = 63.7$ with $\nu = 15$ degrees
of freedom (dof), while $\chi^2 = 10.2$ with $\nu = 14$ dof for the $\alpha =
0.98$ model.

As previously noted, a precise determination of the rise time,
$t_\mathrm{rise}$, of SNe Ia is challenging as there may be a dark phase
following explosion \citep{2014ApJ...784...85P}. Nevertheless, to be
consistent with previous studies (e.g., \citealt{2011MNRAS.416.2607G}), here
we find $t_\mathrm{rise} = 17.91\pm _{0.07}^{0.15}$ d based on our fit for
$t_0$. We caution, however, that $t_0$ corresponds to the time when \abc\ was first detectable by P48 and not the time of explosion.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=3.35in]{early_lc.pdf}
  \caption{
  Best-fit $f \propto t^\alpha$ model to describe the early flux from \abc\ in
  the $g_\mathrm{PTF}$ band. \textit{Top}: The relative flux,
  $f_{g_\mathrm{PTF}}$, shown as green circles, is measured via forced-PSF
  photometry. The model flux, adopting best-fit parameters $\alpha=0.98$ and
  $t_0=-17.91\,\textrm{d}$, is shown as a thick, dashed line. Also shown is
  the best-fit model after fixing $\alpha=2$ (thin, dash-dot line). The inset
  shows the joint distribution of $t_0$ and $\alpha$ for the best-fit
  power-law model. The solid contours represent the $68\%$ and $99.7\%$
  confidence levels. \textit{Bottom}: The observations and models following
  subtraction of the best-fit power-law model, $f_{\alpha = 0.98}$. The
  $t^{0.98}$ model provides a much better fit to the observations than the
  $t^2$ model.}
  \label{fig:early_lc_fit}
\end{figure}

Unlike \abc, the early emission from SN\,2011fe is well fit by a $f \propto
t^2$ model \citep{2011Natur.480..344N}. Thus, the near-linear flux evolution
observed in \abc\ is distinct compared to SN\,2011fe. To our knowledge this
behavior has only been observed in 2 other SNe (2013dy, 2014J;
\citealt{2013ApJ...778L..15Z,2014ApJ...783L..24Z,2015ApJ...799..106G}). Any
model to explain the observations of \abc\ must account for this near-linear
rise in the days after first light.

As a brief aside, we note that simulations presented in
\citet{2017MNRAS.472.2787N} show that SNe Ia explosion models do not evolve as
a power law in time. \citeauthor{2017MNRAS.472.2787N\ demonstrate
that $f \propto t^\alpha$ fits to simulated light curves result in glaring
errors to the estimated explosion times. While caution is advised in
\citet{2017MNRAS.472.2787N}, we note that our primary aim with the power-law
fit is to characterize $\alpha$ for \abc\ compared to SN\,2011fe.}

\subsection{Time of Explosion from the Photospheric Velocity}
\label{sec:early_vel}

The time of explosion $t_\mathrm{exp}$ is not equal to $t_0$ (see
above), thus, \citet{2014ApJ...784...85P} suggest that measurements of the
photospheric velocity can determine $t_\mathrm{exp}$ given that the ejecta
begin expanding from the moment of explosion. Assuming a constant opacity in
the ejecta, \citeauthor{2014ApJ...784...85P} find that the photospheric
velocity evolves as $v_\mathrm{ph}\propto(t-t_\mathrm{exp})^{-0.22}$.
Numerical experiments by \citet{2016ApJ...826...96P} find that the
constant-opacity assumption strongly depends on the amount of $^{56}$Ni mixing
in the SN ejecta. As a result, the adoption of a $t^{-0.22}$ power-law model
may not be valid for all SNe Ia. Nevertheless, we proceed on the assumption
that \abc\ experienced strong $^{56}$Ni mixing (see
\S\ref{sec:Ni_mixing)}, corresponding to the models that are best
approximated as a $t^{-0.22}$ power law. We do this in part to compare with
previous studies, though we caution that the inferred value of $t_{exp}$ is
subject to uncertainties related to ejecta mixing.

While the photospheric velocity is not easy to measure, line velocities of
\ion{Si}{2} or \ion{Ca}{2} can be used as a proxy
\citep{2014ApJ...784...85P,2016ApJ...826..144S}. In the case of \abc, the
\ion{Ca}{2} IR triplet is very weak, likely due to high temperatures in the
ejecta. Thus, we determine the photospheric velocity from the
\ion{Si}{2}\,$\lambda$6355 line. Visual inspection shows no sign of
multi-velocity components of \ion{Si}{2}, and that the
\ion{C}{2}\,$\lambda$6580 line overlaps the red wing of the \ion{Si}{2} line
(see Figures~\ref{fig:spec_seq} and \ref{fig:carbon}). Consequently, we model
the observed spectra between 5900 and 6500\,\AA\ (rest-frame) as the
combination of two Gaussian kernels plus a linear baseline, which accounts for
\ion{Si}{2}, \ion{C}{2} and the continuum, respectively. The expansion
velocity of \ion{Si}{2} is measured by the central wavelength of the
\ion{Si}{2} Gaussian kernel.

\begin{figure*}[!thb]
  \centering
  \includegraphics[width=3.35in]{Chi2.pdf}
  \includegraphics[width=3.35in]{VelocityPlot.pdf}
  \caption{
  Constraints on $t_\mathrm{exp}$ from the velocity evolution of \ion{Si}{2}.
  \textit{Left panel:} the dashed, solid and dash-dotted curves show $\chi^2$
  for fitting power laws with indices $-0.20$, $-0.22$ and $-0.24$,
  respectively. The blue vertical line and the orange shaded region indicate
  $t_0$ and its 95\% confidence interval from Section \ref{sec:lc_fit},
  respectively. \textit{Right panel:} Observed \ion{Si}{2}\,$\lambda$6355
  velocities (blue circles) and the best-fit power-law model with an index of
  $-0.22$ (dashed line). For comparison, the measured velocities of
  \ion{C}{2}\,$\lambda\lambda$6580, 7234 are also shown. Typical uncertainties
  are smaller than the size of the points. \textit{Right inset}: Evolution of
  the pseudo-equivalent width of \ion{Si}{2}\,$\lambda$6355, and
  \ion{C}{2}\,$\lambda\lambda$6580, 7234 in the $\sim$7\,d following 
  explosion.}
  \label{fig:velocity_t_exp}
\end{figure*}

We fit the measured velocities of \ion{Si}{2}\,$\lambda$6355 to the
$v_\mathrm{ph}\propto(t-t_\mathrm{exp})^{-0.22}$ model by minimizing the
$\chi^2$ value and find the best-fit explosion time relative to
$T_{B,\mathrm{max}}$ in the SN rest frame to be $t_\mathrm{exp} = -17.45 \pm
^{0.14}_{0.16}\,\textrm{d}$, where the uncertainties represent the 95\%
confidence interval (Figure \ref{fig:velocity_t_exp}). Following the analysis
in \citet{2014ApJ...784...85P}, we additionally alter the power-law index to
$-0.20$ and $-0.24$ to examine the sensitivity of the result on the assumed
value of $-0.22$. We find that this variation in the power-law index results
in a change of $t_\mathrm{exp}$ of $\approx\pm$0.5\,d (Figure
\ref{fig:velocity_t_exp}). Given the analytical approximation that
$v_\mathrm{ph} \propto t^{-0.22}$, we adopt $t_\mathrm{exp} = -17.5 \pm 0.5 \;
\mathrm{d}$, where the uncertainty reflects possible variations in the
power-lax index (see \citealt{2014ApJ...784...85P}).

Comparing our estimates for $t_\mathrm{exp}$ and $t_0$
(Figure~\ref{fig:velocity_t_exp}), we find that $t_0\lesssim t_\mathrm{exp}$.
Since physical causality requires $t_\mathrm{exp} \le t_0$, we draw the
qualitative conclusion that $t_0\simeq t_\mathrm{exp}$, which is consistent to
within the uncertainties. This derivation of $t_\mathrm{exp}$ relies on the
assumption $v_\mathrm{ph} \propto t^{-0.22}$, which may not be valid for all
SNe Ia.

\subsection{Strong and Short-Lived Carbon Features}
\label{sec:carbon}

The early spectra of \abc\ exhibit unusually strong \ion{C}{2} $\lambda\lambda$6580, 7234 absorption. The evolution of these
spectral features is highlighted in Figure~\ref{fig:carbon}, which shows that
\ion{C}{2}\,$\lambda$6580 is as strong as \ion{Si}{2}\,$\lambda$6355 at $t
\approx -15 \, \mathrm{d}$. The strength of the \ion{C}{2} lines declines with
time, and by $t \approx -10 \, \mathrm{d}$ \ion{C}{2} is no longer detectable.

\begin{figure}[]
  \centering
  \includegraphics[width=3.35in]{CarbonFeature.pdf}
  \caption{
  Evolution of the \ion{C}{2} features observed in the early spectra of \abc.
  The raw spectra are shown in orange, while the solid blue lines show the
  best-fit models (see text for further details). The dark grey vertical lines
  show the measured line centers, and clearly show the decline in the
  photosphere velocity in the $\sim$7~d after explosion (
  \ion{C}{2}\,$\lambda$7234 is not detected in the $-10.9$\,d spectrum). The
  phase of each spectrum relative to $T_{B,\mathrm{max}}$ is labeled.
  }
  \label{fig:carbon}
\end{figure}

Similar to our analysis of the \ion{Si}{2}\,$\lambda$6355 line, we can measure
velocities and pEWs of \ion{C}{2}\,$\lambda\lambda$6580, 7234. We compare the
velocity evolution of \ion{C}{2} to \ion{Si}{2} in the right panel of Figure
\ref{fig:velocity_t_exp}, which also shows the pEWs of these lines. These
measurements confirm the qualitative analysis from Figure~\ref{fig:carbon}:
namely, the strength of \ion{C}{2}\,$\lambda$6580 is similar
to \ion{Si}{2}\,$\lambda$6355 at $t \approx -16 \, \mathrm{d}$ before
decreasing and eventually disappearing around $t = -10$ d.

The detection of \ion{C}{2} in SN Ia spectra is relatively rare as it requires
both unburned carbon, which is likely only present in the outermost layers of
the ejecta, and non-local thermal equilibrium effects in order to excite the
ionized carbon (e.g., \citealt{2007ApJ...654L..53T}). Spectra obtained around
or after $T_{B,_\mathrm{max}}$ rarely show \ion{C}{2} as the photosphere has
receded from the outermost ejecta, while pre-max spectra show evidence for
weak \ion{C}{2} absorption in $\sim$1/4 of all normal SNe Ia (e.g.,
\citealt{2011ApJ...732...30P,2011ApJ...743...27T,2012MNRAS.425.1917S}). While
the sample of SNe Ia with spectra taken within a few days of explosion is
small, SNe\,2013dy and 2017cbv are the only other objects known to have strong
\ion{C}{2} features like \abc\
\citep{2013ApJ...778L..15Z,2017ApJ...845L..11H}. As a counterexample,
SN\,2011fe only exhibited weak \ion{C}{2} features in its first spectra
\citep{2012ApJ...752L..26P}. Thus, models of \abc\ must explain the strong
\ion{C}{2} absorption observed shortly after explosion.

\subsection{Blue Optical Colors Shortly After Explosion}

Multiband observations of \abc\ began $\sim$1.5\,d after discovery, which
allows us to trace its color evolution starting $\sim$1.7\,d after $t_0$. In
Figure~\ref{fig:B-Vcolors} we compare the $(B - V)_0$ color evolution of \abc\
to observations of SN\,2011fe \citep{2016ApJ...820...67Z}. For both SNe the
colors have been corrected for the total inferred reddening along the line of
sight. Interestingly, \abc\ has a nearly flat color evolution up to $t \approx
-10$\,d, while SN\,2011fe initially exhibits red colors before evolving to the
blue.

\begin{figure}[]
  \centering
  \includegraphics[width=3.35in]{16abc_11fe_colors.pdf}
  \caption{$(B - V)_0$ color evolution of \abc\ (squares) 
    compared to SN\,2011fe (stars). The $B$ and $V$ photometry are calibrated
    on the Vega system, and have been corrected for extinction. The data for
    SN\,2011fe are from \citet{2016ApJ...820...67Z}.}
  \label{fig:B-Vcolors}
\end{figure}

Roughly 16\,d prior to $T_{B,_\mathrm{max}}$, the $(B - V)_0$ color of \abc\ is
$\sim$0.5\,mag bluer than SN\,2011fe. Like \abc, SN\,2012cg
\citep{2016ApJ...820...92M} and SN\,2017cbv \citep{2017ApJ...845L..11H}
exhibit $(B - V)_0$ colors that are significantly bluer than SN\,2011fe at
very early epochs. While there are many factors that contribute to the early
optical colors of SNe Ia (see \S\ref{sec:lc_energy} below), early blue colors
are often interpreted as a hallmark of interaction between the SN ejecta and a
binary companion. Nevertheless, it is interesting to note that despite the
blue optical colors, the $\mathrm{UV} - \mathrm{optical}$ colors of \abc,
SN\,2012cg, and SN\,2017cbv are significantly redder at these early epochs
than the $\mathrm{UV} - \mathrm{optical}$ colors of iPTF\,14atg
\citep{2015Natur.521..328C}, the most likely candidate for SN ejecta-companion
interaction.

\section{Modeling the Early Evolution of \abc} \label{sec:lc_energy}

Relative to the nearby, normal SN\,2011fe, we have identified several distinct
characteristics of the early evolution of \abc, including: (i) a near-linear
photometric rise, (ii) a qualitatively short, or possibly absent, dark phase,
assuming $v_\mathrm{ph} \propto t^{-0.22}$, (iii) the presence of strong
\ion{C}{2} absorption, and (iv) blue and nearly constant $(B - V)_0$
color in the week after explosion. While most SNe Ia are powered purely by
radioactive decay, the observed radiation shortly after explosion can also
include contributions from SN shock cooling or the collision of the SN ejecta
with a non-degenerate companion or nearby, unbound material. Here we
consider these scenarios as possible explanations for the early behavior of
\abc.

\subsection{SN Shock Cooling}

The shock breakout of a SN Ia lasts for a fraction of a second due to compact
size of the exploding star. Emission from the subsequent cooling phase may
last for several days, however (e.g., \citealt{2010ApJ...708..598P}).
Following the analysis of \citet{2012ApJ...744L..17B} for SN\,2011fe, we
compare the early-phase $g_\mathrm{PTF}$ light curve of \abc\ with two shock
cooling models \citep{2011ApJ...728...63R, 2010ApJ...708..598P}. From this
analysis, we constrain the \abc\ progenitor radius to be $<1\,\sr$. Our
observations of \abc\ cannot place tight constraints on the size of its
progenitor. Indeed, for a typical WD radius, such as that inferred for
SN\,2011fe ($\lesssim 0.02$--$0.04\,\sr$; \citealt{2012ApJ...744L..17B,
2014ApJ...784...85P}), the expected emission from shock cooling is
$\sim$2\,mag fainter than the P48 $g_\mathrm{PTF}$ detection limit at this
distance. Thus, we conclude that shock cooling does not contribute to the
early emission detected from \abc.

\subsection{SN-Companion collision}
\label{sec:companion}

The detection of emission from the collision of the SN ejecta with a
non-degenerate companion requires a favorable orbital alignment relative to
the line of sight. Thus, from geometric considerations alone the probability
of detecting ejecta-companion interaction is low, $\sim$10\%.
\citet{2010ApJ...708.1025K} calculates that the collision of SN ejecta with a
companion generates thermal emission with a spectrum that peaks in the UV. The
resulting \textit{g}-band emission is expected to be weak.

To examine the possibility of a SN-companion signature in the early light
curve of \abc, we employ the \citet{2010ApJ...708.1025K} model and assume
canonical values for the ejecta mass, $1.4\,\sm$, expansion velocity,
$10^{4}\,\textrm{km}\,\textrm{s}^{-1}$, and a constant opacity,
$0.2\,\textrm{cm}^2\,\textrm{g}^{-1}$. We calculate the expected
$g_\mathrm{PTF}$ brightness of an ejecta-companion collision at the distance
of \abc\ behind a total reddening of $E(B-V) = 0.08 \, \mathrm{mag}$ using the
parameterized equations in \citet{2012ApJ...749...18B}. If we assume the
binary is aligned with the optimal orientation relative to the line of sight,
a binary separation of $a \approx 3 \times 10^{11}\, \mathrm{cm}$ is needed to
explain the initial detection of \abc, as shown in
Figure~\ref{fig:SN-companion}. The minimum binary separation capable of
explaining the observed brightness at the epoch of discovery is $a \approx
10^{11} \, \mathrm{cm}$. Figure~\ref{fig:SN-companion} shows that such models
peak at $g_\mathrm{PTF} \approx 21.5 \, \mathrm{mag}$, provided that
$t_\mathrm{exp} \approx t_0 - 0.3 \, \mathrm{d}$. These models do not,
however, match the $g_\mathrm{PTF}$ evolution for $t > t_0 + 0.5 \,
\mathrm{d}$ (though it is possible that the SN photosphere dominates the
companion-interaction signature at this phase). While they are otherwise
compatible with the observations, we do not favor the above models as the
explanation for the early flux from \abc\ because they do not explain the
2\,mag rise in the $\sim$24 hr after discovery.

\begin{figure}[!thb]
  \centering
  \includegraphics[width=3.35in]{companion_lc.pdf}
  \caption{Comparison of SN ejecta-companion interaction models 
  with early observations of \abc. $g_\mathrm{PTF}$ 
  detections and 3$\sigma$ upper limits are shown 
  as green circles and downward arrows, respectively. 
  \textit{Swift}/$UVW1$ observations are shown as magenta hexagons.
  The dashed, solid, and dash-dot lines show the expected flux 
  for companion-interaction models in the $g_\mathrm{PTF}$ (green) 
  and $UVW1$ (magenta) filters. The models have been adjusted to 
  account for the distance and reddening towards \abc. Each 
  model features a different companion semi-major axis, $a$, and 
  time of explosion, $t_\mathrm{exp}$, as labeled in the 
  legend. While models with $a \gtrsim 10^{12} \, \mathrm{cm}$ can 
  explain the early optical rise, they greatly 
  overpredict the UV flux. Models with $a \approx 10^{11} \, 
  \mathrm{cm}$ can explain the initial detection \abc, but they 
  fail to replicate the $\sim$2\,mag rise in the $\sim$24\,hr 
  after explosion.
  }
  \label{fig:SN-companion}
\end{figure}

Figure~\ref{fig:SN-companion} additionally shows that a companion at $a
\approx 18 \times 10^{11} \, \mathrm{cm}$ provides a good match to the initial
optical rise, if $t_\mathrm{exp} \approx t_0 + 0.1 \; \mathrm{d}$. Models with
$a \gtrsim 10^{12} \, \mathrm{cm}$, which can explain the initial
$g_\mathrm{PTF}$ rise, significantly overpredict the observed UV flux,
however. There is no choice of $a$ capable of replicating the early rise of
\abc\ without also overpredicting the observed UV flux.

The challenges associated with each of the previously considered models lead
us to conclude that the early evolution of \abc\ cannot be explained via
ejecta-companion interaction. We cannot, however, exclude the presence of a
red-giant, or other non-degenerate, companion as our calculations have assumed
that the binary is aligned with the optimal geometry relative to the line of
sight. If the geometry is not favorable, then it is possible that signatures
from interaction with a companion are not visible.

\subsection{Sub-Chandrasekhar Detonations and Pure Deflagrations}

In \citet{2017MNRAS.472.2787N} the early photometric evolution of Type
Ia SNe is explored via a variety of explosion models and detailed
radiative-transfer calculations. Specifically, \citet{2017MNRAS.472.2787N}
examine two Chandrasekhar-mass ($M_\mathrm{Ch}$) explosions and compare their
evolution to sub-Chandrasekhar detonations and pure deflagrations. The
$M_\mathrm{Ch}$ explosions include the ``W7'' carbon-deflagration model of
\citet{1984ApJ...286..644N} and the ``N100'' delayed-detonation model from
\citet{2013MNRAS.429.1156S}. The sub-Chandrasekhar models include: a violent
WD-WD merger, which triggers a carbon detonation in the more massive WD, a
centrally ignited sub-Chandrasehkar detonation, and a sub-Chandrasehkar double
detonation explosion, in which a He-surface-layer detonation triggers a carbon
detonation in the core. \citet{2017MNRAS.472.2787N} note that, of these last
two sub-Chandrasehkar models, the later provides the more realistic scenario.
Finally, \citeauthor{2017MNRAS.472.2787N} also examine the ``N5def'' and
``N1600Cdef'' pure deflagration explosions from \citet{2014MNRAS.438.1762F}.
The ``N5def'' model is particularly unique in that the explosion does not
fully unbind the WD, meaning, unlike typical Type Ia SNe, a remnant remains.

\begin{figure}[]
  \centering
  \includegraphics[width=3.35in]{compare_Noebauer.pdf}
  \caption{
  %
  Comparison of the models from \citet{2017MNRAS.472.2787N to the
  \abc\ $V$-band light curve (\textit{top}) and the $(B-V)_0$ color curve
  (\textit{bottom}). The light curve has been corrected for the distance
  modulus to \abc\ (\S\ref{sec:host}), while both the light curve and color
  curve have been corrected for reddening. To guide the eye,
  $g_\mathrm{PTF}$ observations are shown, though we caution
  that \abc\ may exhibit significant color evolution at this phase in which
  case $g_\mathrm{PTF}$ would be a poor proxy for $V$. Each model light curve
  is translated to match the LCO observations $\sim$1.7\,d after $t_0$ given
  the uncertain time of explosion. Translational offsets are listed in the
  bottom panel legend, which shows the models in order of decreasing
  $^{56}$Ni from top to bottom, then left to right. The inset in
  the top panel shows the residuals relative to the W7 model.} }
  %
  \label{fig:noebauer} 
\end{figure}

In Figure~\ref{fig:noebauer} we compare photometric observations of
\abc\ to the models presented in
\citet{2017MNRAS.472.2787N}.\footnote{Available here:
\url{https://hesma.h-its.org/}.} Interestingly, the sub-Chandrasekhar double
detonation model (subChDoubleDet) replicates the early wiggle in the $V$-band
light curve. However, this match requires an explosion \textit{after} our
initial detection of \abc, and predicts extreme color evolution that is not
observed. Thus, the subChDoubleDet model is incompatible with the
observations. The sub-Chandrasekhar detonation (subChDet) provides a better
match to the observations, though this model is not favored as a particularly
realistic scenario (see above). Of the sub-Chandrasekhar models the violent
merger model (Merger) provides the best match to the observations including
the early rise and color evolution. In detail, however, this model does not
match the early wiggles in the light curve, has consistently redder colors
than \abc, and requires $t_\mathrm{exp} \approx 2.3$\,d prior to $t_0$. As
such, we postulate that \abc\ is not the result of a violent WD-WD merger.

For clarity, of the two pure deflagration models only N5def is shown,
however, the evolution of N1600Cdef is very similar. While the pure
deflagration models produce the bluest colors at early times, they are
underluminous at times $> t_0 + 4\,\mathrm{d}$ and already rapidly evolving
towards the red at $\sim$$t_0 + 7$ (\abc\ exhibits a nearly constant $(B-V)_0$
color for $\sim$19\,d after $t_0$). Thus, we conclude that \abc\ is not
compatible with pure deflagrations.

Of the $M_\mathrm{Ch}$ models, the W7 model better matches the
observations as the N100 model features a faster rise and higher luminosity
than what is observed. We explore delayed detonation models in further detail
below.}

\subsection{Interaction with Nearby, Unbound Material}

To model SN\,2011fe, \citet{2014MNRAS.441..532D} examined
pulsational delayed-detonation (PDD) models as an explanation for some SNe
Ia. Briefly, PDD models differ from ``standard'' delayed detonation (DD)
models in that the expansion of the WD during the initial deflagration
phase leads to the release of unbound material. Following this pulsation,
the bound material contracts eventually triggering a subsequent
detonation.\footnote{\citet{2014MNRAS.441..532D} note that the deflagration
and detonation in their PDD models are artificially triggered.} An
important consequence of this progression for PDD models is that the
unbound material expands and avoids burning, unlike DD models that
typically leave no unburnt material. This results in significantly more
carbon in the outer layers of the SN ejecta \citep{2014MNRAS.441..532D}.

\citet{2014MNRAS.441..532D} find that DD models are universally faint and red at early times, $\sim$24--48\,hr after explosion, while the PDD models exhibit a faster rise and bluer colors. Briefly, this occurs in the PDD scenario because the collision with the unbound material surrounding the WD heats the outer layers of the SN ejecta. Importantly, the PDD models are nearly indistinguishable from DD models around and post peak.

\begin{figure}[]
  \centering
  \includegraphics[width=3.35in]{compare_Dessart.pdf}
  \caption{
  %
  Same as Figure~\ref{fig:noebauer}, featuring the $g$-band light
  curve and the DD and PDD models from \citet{2014MNRAS.441..532D}. The bottom
  panel legend lists the models in order of decreasing $M_\mathrm{Ni}$ from
  top to bottom. DD models (labeled as DDC to match the nomenclature of
  \citealt{2014MNRAS.441..532D}) are shown as dash-dot lines, while PDD models
  (labeled as PDDEL) are shown as dashed lines. The inset in the top panel
  shows the residuals relative to the PDDEL7n model. The PDD models provide a
  better match to the observations.}
  %
  \label{fig:dessart} 
\end{figure}

In Figure~\ref{fig:dessart} we compare photometric observations of
\abc\ to the high-$^{56}$Ni yield DD and PDD models presented in
\citet{2014MNRAS.441..532D}.\footnote{Available here:
\url{https://www-n.oca.eu/supernova/snia/snia_ddc_pddel.html}.} DD and PDD
models with $M_\mathrm{Ni}/M_\odot \la 0.7$ and 0.5, respectively, fail to
match the early luminosity and blue colors of \abc. The PDD models provide a
better match to the observations than the DD models. The PDD models evolve
more rapidly towards blue colors and provide a better match to the $g$-band
flux in the days after explosion. Furthermore, unlike the DD models, the PDD
models exhibit strong \ion{C}{2} lines that gradually disappear in the
$\sim$week after explosion \citep{2014MNRAS.441..532D}. In detail, the early
wiggles in the \abc\ light curve are not matched by the PDD models, which
exhibit more smooth variations. Furthermore, while the focus of this study is
the early evolution of \abc, the PDD models evolve more rapidly to the red
post-peak than the observations. Nevertheless, the PDD models presented in
\citet{2014MNRAS.441..532D} provide several attractive explanations for the
unusual features in the early behavior of \abc. Small adjustments to the PDD
models (e.g., additional $^{56}$Ni mixing, which
\citeauthor{2014MNRAS.441..532D} only explore for DD models) may better match
\abc.

\subsection{Strong $^{56}Ni$ Mixing in the SN Ejecta}
\label{sec:Ni_mixing}

Having examined other possibilities we now consider whether the early evolution of \abc\ can be explained simply by invoking strong mixing in the 
SN ejecta. Strong mixing leads to a faster initial rise as well as a more rapid evolution towards blue colors.

\begin{figure}[]
  \centering
  \includegraphics[width=3.35in]{compare_Piro.pdf}
  \caption{
  %
  Same as Figure~\ref{fig:noebauer}, featuring the $g$-band light
  curve and models from \citet{2016ApJ...826...96P}. The amount of $^{56}$Ni
  mixing in the SN ejecta increases from the short-dash, light lines to the
  long-dash, dark lines. Unlike Figure~\ref{fig:dessart}, each model features
  the same $M_\mathrm{Ni}$, while the model names reflect the boxcar widths
  used to approximate the effects of mixing in the ejecta (see
  \citealt{2016ApJ...826...96P}). The top panel inset shows the residuals
  relative to the $0.25\,M_\odot$ model. The observations are best matched by
  the $0.25\,M_\odot$ model, i.e. the model with the most significant mixing.}
  %
  \label{fig:piro}
\end{figure}

Figure~\ref{fig:piro} compares the models from \citet{2016ApJ...826...96P} to
\abc. The \citeauthor{2016ApJ...826...96P} models employ a piston driven
explosion to explode a single WD progenitor model. As the piston explosion
does not result in any nucleosynthesis, the distribution of $^{56}$Ni in the
ejecta must be prescribed by hand, which enables a study of the effects of
mixing on the resulting SN emission. Each model employs a fixed $0.5\,\sm$ of
$^{56}$Ni that has been distributed throughout the ejecta via boxcar averaging
(see their Figure~1). The resulting light curves are synthesized using the
SuperNova Explosion Code (\texttt{SNEC}; \citealt{2015ApJ...814...63M}), as
shown in Figure~\ref{fig:piro}. Broadly speaking, the results can be
summarized as follows: SNe with strong mixing exhibit a rapid rise and quickly
develop blue colors, whereas models where the $^{56}$Ni is confined to the
innermost layers of the ejecta remain faint for days after explosion and
feature a gradual color evolution from the red to the blue. The model
with the strongest mixing (dark, long-dashed line in Figure~\ref{fig:piro)}
best matches the observations of \abc. This is the only model we have found
that exhibits a flat $(B-V)_0$ color evolution in the days after explosion,
however, in detail this model is too red relative to the observations of \abc.

While the models from \citet{2016ApJ...826...96P} provide a good match to the
optical photometric evolution of \abc, they consistently overpredict the flux
in the UV. They also overpredict the photospheric velocity of \abc\
by $\sim$2--3,000$\; \mathrm{km \; s}^{-1}$. Furthermore, the simple
grey opacities in SNEC likely produce a faster rise and bluer colors than the
more detailed treatments employed in \citet{2014MNRAS.441..532D} and
\citet{2017MNRAS.472.2787N}.

Both the PDD models and ejecta-mixing models show discrepancies with some
early observations of \abc. Nevertheless, we conclude that one, or both, of
these scenarios, which feature qualitatively similar predictions, is the most
likely explanation for \abc. Indeed, it may be the case that the typical
sequence of photometric and spectroscopic observations of young SNe Ia can
never distinguish between these two possibilities \citep{2017MNRAS.472.2787N}.

\section{The Emerging Sample of Young \sneia}

The proliferation of high-cadence, time-domain surveys has led to several SNe
Ia being discovered within $\sim$2\,d of first light in the past $\sim$decade.
Observations probing the early evolution of these SNe allows us to place
unique constraints on their progenitor systems and the corresponding explosion
physics. This has revealed diversity in the earliest epochs after explosion,
and that commonly used SNe Ia templates do not match observations at these
phases (e.g., \citealt{2012ApJ...744...38F}).

In \S\ref{sec:first_light} we compared our early observations of \abc\ to
SN\,2011fe, which has the most comprehensive observations of the young SN Ia
sample. Given its normal spectroscopic and photometric evolution, SN\,2011fe
has been adopted as a standard for the early evolution of SNe Ia in many
studies. While a detailed quantitative analysis is beyond the scope of this
study, a qualitative examination of very young SNe Ia that otherwise exhibit
normal spectra and evolution at and post peak\footnote{This definition
excludes iPTF\,14atg, which was shown to be subluminous with SN\,2002es-like
spectra \citep{2015Natur.521..328C}.} reveals considerable diversity. In other
words, at early times SN\,2011fe may not be the norm.

For SN\,2011fe the initial rise is well described by a $t^2$ power law, the
$(B - V)_0$ colors evolve from the red to the blue in the $\sim$week after
explosion, and the \ion{C}{2} present in the initial spectra is weak
\citep{2011Natur.480..344N,2016ApJ...820...67Z,2012ApJ...752L..26P}. In
contrast, \abc\ exhibits a near linear rise in flux, the $(B - V)_0$ colors
are blue and roughly constant, and the \ion{C}{2} absorption is strong.
Examining just these 3 qualitative features, SN\,2009ig is well matched to
SN\,2011fe \citep{2012ApJ...744...38F}, while SN\,2013dy
\citep{2013ApJ...778L..15Z}, SN\,2017cbv \citep{2017ApJ...845L..11H}, and
\abc\ all bear a striking resemblance. SN\,2012cg, on the other hand, is
intermediate to these two groups with weak C II and a relatively shallow early
rise, like SN\,2011fe, but blue $(B - V)_0$ colors, like \abc\
\citep{2012ApJ...756L...7S,2016ApJ...820...92M}. SN\,2014J is intermediate in
the other direction in that it exhibits a near linear rise
\citep{2014ApJ...783L..24Z,2015ApJ...799..106G}, but the color evolution is
very similar to SN\,2011fe \citep{2014ApJ...788L..21A}.\footnote{\ion{C}{2} is
not detected in the spectra of SN\,2014J
\citep{2014ApJ...784L..12G,2014ApJ...783L..24Z}, though the earliest spectra
of SN\,2014J were obtained at a much later phase than the other SNe discussed
here. \citet{2015ApJ...798...39M} find evidence for \ion{C}{1} in the NIR
spectra of SN\,2014J, but this detection cannot constrain the relative
strength of \ion{C}{2} and \ion{Si}{2} shortly after explosion.} That these
early observations cannot be easily separated into two distinct groups
suggests that it is unlikely that a single physical mechanism drives the
diversity of SNe Ia at early times.

In the case of SN\,2012cg and SN\,2017cbv it has been argued that the early
blue optical colors are indicative of interaction between the SN ejecta and a
binary companion \citep{2016ApJ...820...92M,2017ApJ...845L..11H}. SN\,2017cbv
is particularly remarkable in that the observations presented in
\citet{2017ApJ...845L..11H} show a clearly resolved bump in the $U$, $B$, and
$g'$ bands in the $\sim$5\,d after explosion. In
\citeauthor{2017ApJ...845L..11H} it is found that the bump can be explained
via the combination of ejecta-companion interaction and the normal evolution
of a SN Ia. A challenge for this model, similar to \abc\ (see
\S\ref{sec:companion}), is that it significantly overpredicts the UV
brightness of the SN compared to what is observed. Indeed, in the case of
SNe\,2012cg, 2017cbv, and \abc\ the $\mathrm{UV} - \mathrm{optical}$ colors
are significantly redder than those observed in iPTF\,14atg. It is argued in
\citet{2017ApJ...845L..11H} that several model assumptions, including: (i)
ideal blackbody emission, (ii) a constant opacity, (iii) a simple power-law
density profile for the ejecta, and (iv) spherical symmetry, may be incorrect,
which could reconcile the discrepancy with the UV observations. Above, we
argued for interaction with diffuse, unbound material and strong
$^{56}$Ni mixing as a possible explanation for \abc, and indeed
\citet{2017ApJ...845L..11H} consider these possibilities for SN\,2017cbv as
well. Separately, several arguments against companion interaction for
SN\,2012cg are presented in \citet{2016arXiv161007601S}.

Ultimately, there are arguments in favor of and against each of the
possibilities to model the early emission from SNe Ia. Moving
forward, more detailed models and simulations are needed to properly explain
the observed diversity. No matter the correct explanation for the early
behavior of SNe\,2013dy, 2017cbv, and \abc, the strong similarities between
these events suggests that they may reflect a common physical origin.

\section{Conclusion}
\label{sec:conclusion}

We have presented observations of the extraordinarily early discovery of the 
normal Type Ia supernova \abc. Our fast-response follow-up 
campaign allowed us to draw the following conclusions:

\begin{itemize}

    \item Extrapolation of the early light curve shows that the initial
    detection of \abc\ occurred only $0.15\pm_{0.07}^{0.15}$~d after the time
    of first light, $t_0$.

    \item We find no evidence for detectable signatures of SN shock cooling or
    the collision of the SN ejecta with a non-degnerate binary companion.

    \item Assuming that $v_\mathrm{ph} \propto t^{-0.22}$, then
    $t_\mathrm{exp} \approx t_0$. A short dark phase, as this implies, is
    likely the result of either strong $^{56}$Ni mixing or interaction of the
    SN ejecta with nearby, unbound material.

    \item The strong and short-lived carbon features seen in the earliest
    spectra of \abc\ can only be explained if there is incomplete burning. The
    pulsational delayed detonation models presented in
    \citet{2014MNRAS.441..532D} produce \ion{C}{2} absorption that is as
    strong as \ion{Si}{2} at very early phases.

    \item In contrast to SN\,2011fe, $(B - V)_0$ is $\sim$0.5 mag bluer for
    \abc\ at $t \approx -16$\,d. Furthermore, the $(B - V)_0$ colors of \abc\
    show no evolution over the first $\sim$7\,d of observations.

    \item Finally, we show that the early light curve evolution and colors of
    \abc\ are best matched by the pulsational delayed detonation
    models of \citet{2014MNRAS.441..532D} and the ejecta-mixing models of
    \citet{2016ApJ...826...96P}.

\end{itemize}
%
Taken together, these observations suggest that the early emission
from \abc\ is due to either the collision of the SN with nearby, unbound
material and/or that there is significant mixing of $^{56}$Ni in the SN
ejecta. The PDD models from \citet{2014MNRAS.441..532D} are particularly
attractive for explaining \abc, because they produce strong
\ion{C}{2} absorption at early times. In the future, it would be useful to
investigate more detailed PDD models that incorporate strong $^{56}$Ni mixing
to see if they better replicate the observations of \abc\ as it is otherwise
difficult to distinguish between these two scenarios.

Extremely early observations of young SNe provide a ``smoking
gun'' to probe the mixing level in the ejecta, which, in turn, is 
a result of the explosion mechanism. Wide-field, high-cadence surveys, such as the Zwicky Transient Facility \citep{2016PASP..128h4501B} and ATLAS \citep{2011PASP..123...58T,2013RSPTA.37120269T}, will discover a large number of very young SNe over the next few years, allowing us to extend our studies beyond single objects. While the sample of extremely young SNe Ia will grow by more than an order of magnitude, the detection of shock breakout cooling and ejecta-companion interaction will prove challenging. Given the diminutive size of WDs, the thermal emission following shock breakout can only be detected to $\sim 10\,\mathrm{Mpc}$ on 1-m class telescopes. Furthermore, only $\sim$10\% of single-degenerate progenitors are expected to give rise to detectable emission following the collision of the SN ejecta with the binary companion \citep{2010ApJ...708.1025K}. Despite these limitations, this study of \abc\ shows that the early detection of SNe Ia may probe explosion physics by measuring the amount of mixing in the SN ejecta. Moving forward, a large sample of such objects will enable strict constraints on the proposed explosion mechanisms for Type Ia SNe.

Finally, we close by emphasizing the importance of fast-response photometric and spectroscopic follow-up campaigns. Without the early recognition of the youth of this SN, and the associated follow-up, much of the analysis presented herein would not have been possible. The ability to trigger such observations is essential to improve our physical understanding of Type Ia SNe.

\acknowledgements

This study has benefited from the suggestions of an anonymous
referee. We are deeply grateful to R.~C.~Thomas for indulging a slew of
questions regarding \ion{C}{2} in SNe. Similarly, it is our pleasure to buy a
beer for R.~Amanullah and U.~Feindt for discussions regarding SALT2.

This paper utilizes the LCO infrastructure for rapid and regular monitoring of
SNe, and we thank S.~Valenti and I.~Arcavi for their development efforts.
Figure~\ref{fig:branch_vel} would not have been possible without
S.~Blondin generously sharing the data from \citet{2012AJ....143..126B}.

Much of the analysis presented herein would not have been possible without the
help of several observers. We thank M.~West for taking the first spectrum of
\abc\ as a ToO on the DCT. We also thank P.~GuhaThakurta, E.~C.~Cunningham,
K.~A.~Plant, H.~Jang, and J.~Torres for executing a Keck ToO as part of the
UC/Caltech partnership, and also the Gemini service observers for executing
our ToO observations. Additionally, J.~Cohen, N.~Suzuki, V.~Ravi, R.~Walters,
A.~Ho, H.~Vedanthamand, K.~De, and L.~Yan helped obtain data for this paper.

AAM is funded by the Large Synoptic Survey Telescope Corporation in support of
the Data Science Fellowship Program. YC thanks support from a postdocal
fellowship at the eScience Institute, University of Washington.

DAH, CM, and GH are supported by NSF-1313484.

The Intermediate Palomar Transient Factory project is a scientific collaboration among the California Institute of Technology, Los Alamos National Laboratory, the University of Wisconsin, Milwaukee, the Oskar Klein Center, the Weizmann Institute of Science, the TANGO Program of the University System of Taiwan, and the Kavli Institute for the Physics and Mathematics of the Universe. This work was supported by the GROWTH project funded by the National Science Foundation under Grant No 1545949. Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the NASA. This work makes use of observations from the LCO network. These results made use of the Discovery Channel Telescope at Lowell Observatory. Lowell is a private, non-profit institution dedicated to astrophysical research and public appreciation of astronomy and operates the DCT in partnership with Boston University, the University of Maryland, the University of Toledo, Northern Arizona University and Yale University. The upgrade of the DeVeny optical spectrograph has been funded by a generous grant from John and Ginger Giovale. Based on observations made with the Nordic Optical Telescope, operated by the Nordic Optical Telescope Scientific Association at the Observatorio del Roque de los Muchachos, La Palma, Spain, of the Instituto de Astrof\'isica de Canarias.

\facility{DCT, Gemini:Gillett, Hale, Keck:I, Keck:II, LCOGT, PO:1.2\,m, PO:1.5\,m, NOT, VLT, Swift, OANSPM:HJT}

\software{\texttt{PTFIDE} \citep{2017PASP..129a4002M}, \texttt{FPipe}, \texttt{SALT2}, \texttt{SNID}, \texttt{lcogtsnpipe}, \texttt{UVOTSOURCE}}.

\appendix

\section{Photometric Light Curves}

The full photometric light curves of \abc\ are shown in Figure~\ref{fig:ugly_LC}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=7in]{lightcurve2.pdf}
  \caption{
  %
  UV, optical, and NIR light curves for \abc. \textit{Left}: $BVgri$ light
  curves from the P48, P60, LCO-1m, \textit{Swift}, and SPM-1.5m telescopes.
  The solid lines represent the best-fit model from SALT2 (see
  \S\ref{sec:classification}). \textit{Right}: UV and NIR light curves from
  the \textit{Swift} and SPM-1.5m telescopes, respectively. For both panels
  each light curve is represented with a different color, while the different
  symbols correspond to different instruments, as detailed in the legends. The
  legends also list offsets applied to each light curve. Photometry is shown
  in the AB mag system, with the exception of the $BV$ bands, which are shown
  in the Vega system. These light curves have not been corrected for
  line-of-sight extinction.}
  %
  \label{fig:ugly_LC}
\end{figure}

\bibliographystyle{aasjournal}
\bibliography{ref}

\end{document}